---
title: "RWorksheet_Suero#4b"
author: "Jiruel Suero"
date: "2023-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



1.Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1.

It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r name, echo = TRUE}
vectorA <- c(1,2,3,4,5)

matrix_A <- matrix(0,nrow = 5, ncol =5)

for (j in 1:5)
  for (k in 1:5)
  {
    matrix_A[j,k] <- abs (vectorA[j] - vectorA[k])
  }

matrix_A
```


2.Print the string "*" using for() function. The output should be the same as shown
in Figure 2.

```{r name, echo = TRUE}
for (j in 1:5) {
  cat(paste0("\"", rep("*", j), "\""), "\n")
}
```
 
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r name, echo = TRUE}
userinput <- as.integer(readline("Enter starting number for Fibonacci sequence: "))

x <- userinput
y <- 0


cat("Fibonacci sequence starting from", userinput, ":\n")

repeat {
  
  nextNum <- x + y
  
  if (nextNum > 500){
    break
  }
  cat(nextNum, " ")
  x <- y
  y <- nextNum
  
}

```
4. Import the dataset as shown in Figure 1 you have created previously.

4a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? 
Show your codes and its result.

```{r name, echo = TRUE}
write.csv(HOUSEHOLD_DATA, file = "HOUSEHOLD_DATA.csv")
imported<- read.csv("HOUSEHOLD_DATA.csv")
imported
importedx<-head(HOUSEHOLD_DATA, 6)
importedx


```
4b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? 
Write the R scripts and its output.

```{r name, echo = TRUE}
males <- imported[imported$Gender == "M",]
males

females <- imported[imported$Gender == "F",]
females

numofMale <- nrow(males)
numofMale

numofFem <- nrow(females)
numofFem


```

4c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors.  Write the R scripts and its result

```{r name, echo = TRUE}
totalMaleFemale <- table(imported$Gender)
barplot(totalMaleFemale,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("green", "purple"),
        legend.text = rownames(totalMaleFemale),
        beside = TRUE)


```
5. The monthly income of Dela Cruz family was spent on the following:

```{r name, echo = TRUE}

spending_data <- data.frame(
  Category = c("Food", "Electricity", "Savings", "Miscellaneous"),
  Value = c(60, 10, 5, 25)
)

spending_data$Percentage <- spending_data$Value / sum(spending_data$Value) * 100

colors <- c("blue", "red", "yellow", "black")

pie(spending_data$Value,
    labels = paste(spending_data$Category, "(",spending_data$Percentage,"%)"),
    col = colors,
    main = "Monthly Income Spending of Dela Cruz Family")

legend("topright", spending_data$Category, fill = colors)



```
6a. Check for the structure of the data set using the str() function. Describe what you have seen in the output.

```{r name, echo = TRUE}
data(iris)
str(iris)


```
6b.Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r name, echo = TRUE}
meanIris <- c(mean(iris$Sepal.Length), mean(iris$Sepal.Width), mean(iris$Petal.Length), mean(iris$Petal.Width))
meanIris




meanSepalLength <- mean(iris$Sepal.Length)
meanSepalWidth <- mean(iris$Sepal.Width)
meanPetalLength <- mean(iris$Petal.Length)
meanPetalWidth <- mean(iris$Petal.Width)

meanIris <- data.frame(Sepal_Length = meanSepalLength,
                        Sepal_Width = meanSepalWidth,
                        Petal_Length = meanPetalLength,
                        Petal_Width = meanPetalWidth)
meanIris

```
6c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.

```{r name, echo = TRUE}
species_count <- table(iris$Species)
pie(species_count, labels = species_count, col = rainbow(length(species_count)), main = "Species Distribution")
legend("topright", names(species_count), fill = rainbow(length(species_count)))
```


6d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.

```{r name, echo = TRUE}
setosaSubset <- subset(iris, Species == "setosa")
versicolorSubset <- subset(iris, Species == "versicolor")
virginicaSubset <- subset(iris, Species == "virginica")


tail(setosaSubset, 6)
tail(versicolorSubset, 6)
tail(virginicaSubset, 6)
```
6e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.

```{r name, echo = TRUE}
iris$Species <- as.factor(iris$Species)

# Convert the "Species" column to a factor
iris$Species <- as.factor(iris$Species)

# Create a scatterplot
plot(
  Sepal.Length ~ Sepal.Width, 
  data = iris,
  pch = as.integer(iris$Species),  # Use different pch symbols for each species
  col = as.integer(iris$Species),  # Use different colors for each species
  xlab = "Sepal Length",
  ylab = "Sepal Width",
  main = "Iris Dataset",
  sub = "Sepal width and length"
)

# Add a legend
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 1:3, cex = 0.8, title = "Species")

```
